LCA-GB

### 2022-06-27 LCA-GB.Rmd
combine r and python code in RMarkdown

# Data preparation/sampling
```{r}
# Rasterising LANDMAP data
VS <- read_sf("http://lle.gov.wales/catalogue/item/LandmapVisualSensory.json") %>%
  st_make_valid() %>%
  mutate(ScenicQuality = factor(VS_46, levels = c("Low", "Moderate", "High", "Outstanding")),
             Integrity = factor(VS_47, levels = c("Low", "Moderate", "High", "Outstanding")),
             Character = factor(VS_48, levels = c("Low", "Moderate", "High", "Outstanding")),
                Rarity = factor(VS_49, levels = c("Low", "Moderate", "High", "Outstanding"))) %>%
  dplyr::select(ScenicQuality, Integrity, Character, Rarity) %>%
  fasterize(raster(extent(.), resolution = 50, crs = 27700),
            field = "ScenicQuality")

@@

p9 <- bespoke_plot(observation = hex.5km, 
                   formula = Rug ~ SQ,
                   title = "Relationship between Scenic Quality and Ruggedness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Ruggedness of terrain")
p10 <- bespoke_plot(observation = hex.2km, 
                   formula = Rug ~ SQ,
                   title = "Relationship between Scenic Quality and Ruggedness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Ruggedness of terrain")
```
# Ordinal regression
#
```{r}
##### 3.2. Include wildness components as additional predictors
model1 = polr(SQ ~ scale(Sce), data = as(hex.5km, "Spatial"), Hess = TRUE)
model2 = polr(SQ ~ scale(Sce) + scale(Abs), data = as(hex.5km, "Spatial"), Hess = TRUE)
model3 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat), data = as(hex.5km, "Spatial"), Hess = TRUE)
model4 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem), as(hex.5km, "Spatial"), Hess = TRUE)
model5 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem) + scale(Rug), as(hex.5km, "Spatial"), Hess = TRUE)
model6 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rug), as(hex.5km, "Spatial"), Hess = TRUE)

table(model1)
table(model2)
table(model3)
table(model4)
table(model5)
table(model6)
```
```{r}
##### 3.2. Include wildness components as additional predictors
model1 = polr(SQ ~ scale(Sce), data = as(hex.2km, "Spatial"), Hess = TRUE)
model2 = polr(SQ ~ scale(Sce) + scale(Abs), data = as(hex.2km, "Spatial"), Hess = TRUE)
model3 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat), data = as(hex.2km, "Spatial"), Hess = TRUE)
model4 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem), as(hex.2km, "Spatial"), Hess = TRUE)
model5 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem) + scale(Rug), as(hex.2km, "Spatial"), Hess = TRUE)
model6 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rug), as(hex.2km, "Spatial"), Hess = TRUE)

@@

# GLM
```{r}
##### 3.1. Global Model: Generalised Linear Model
# hex.5km %>%
#   drop_na() %>%
#   as("Spatial") %>%
#   glm(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem) + scale(Rug),
#       data = .,
#       family = "poisson") %>%
#   summary()
# 
# R2 = 1 - glm$deviance/glm$null.deviance
```

# Machine learning - Gradient Boosted 
```{r}
# GBM hyper-parameters tuning
# From kaggle: https://www.sciencedirect.com/science/article/pii/S0378778817320844
##### Split data into training and validation dataset
set.seed(2046)
TrainIndex <- hex.2km %>% drop_na() %>% .$SQ %>% createDataPartition(p = .8, list = FALSE, times = 1)
Train <- hex.2km[TrainIndex, ]
Test  <- hex.2km[-TrainIndex, ]

@@

# Reference
# https://github.com/topepo/caret/issues/389
```
```{r}
# Model evaluation
#predicted = predict(xgb_tune, x_test)
# Predict outcomes with the test data
Test %>% dplyr::select(Sce, Abs, Nat, Rem, Rug) %>%
  st_drop_geometry() %>% 
  as.matrix() %>% 
  xgb.DMatrix() %>%
  predict(xgb_tune, ., reshape = T) %>%
  plyr::mapvalues(from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")) %>%
  data.frame(predict = .) %>%
  data.frame(actual = plyr::mapvalues(Test$SQ, from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")),
             predict = .) ->
  result

@@

hex.2k$SQ.pred <-
  hex.2k %>%
  st_drop_geometry() %>%
  dplyr::select(Sce, Abs, Nat, Rem, Rug) %>%
  predict(xgb_tune, ., reshape = T) %>%
  plyr::mapvalues(from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")) %>%
  as.factor()
```
# Visualisation
```{r}
theme_map <- function(...) {
  theme_minimal() +
    theme(
      text = element_text(family = "Arial", color = "#22211d"), #"Ubuntu Regular"
      axis.line = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      # panel.grid.minor = element_line(color = "#ebebe5", size = 0.2),
      panel.grid.major = element_blank(), #element_line(color = "#ebebe5", size = 0.2),
      panel.grid.minor = element_blank(),
      plot.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA), 
      panel.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA), 
      legend.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA),
      panel.border = element_blank(),
      ...
    )
}

>>>>

# Data preparation/sampling

```{r}
# Rasterising LANDMAP data
VS <- read_sf("http://lle.gov.wales/catalogue/item/LandmapVisualSensory.json") %>%
  st_make_valid() %>%
  mutate(ScenicQuality = factor(VS_46, levels = c("Low", "Moderate", "High", "Outstanding")),
             Integrity = factor(VS_47, levels = c("Low", "Moderate", "High", "Outstanding")),
             Character = factor(VS_48, levels = c("Low", "Moderate", "High", "Outstanding")),
                Rarity = factor(VS_49, levels = c("Low", "Moderate", "High", "Outstanding"))) %>%
  dplyr::select(ScenicQuality, Integrity, Character, Rarity) %>%
  fasterize(raster(extent(.), resolution = 50, crs = 27700),
            field = "ScenicQuality")

@@

p9 <- bespoke_plot(observation = hex.5km, 
                   formula = Rug ~ SQ,
                   title = "Relationship between Scenic Quality and Ruggedness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Ruggedness of terrain")
p10 <- bespoke_plot(observation = hex.2km, 
                   formula = Rug ~ SQ,
                   title = "Relationship between Scenic Quality and Ruggedness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Ruggedness of terrain")
```

# Ordinal regression

#

```{r}
##### 3.2. Include wildness components as additional predictors
model1 = polr(SQ ~ scale(Sce), data = as(hex.5km, "Spatial"), Hess = TRUE)
model2 = polr(SQ ~ scale(Sce) + scale(Abs), data = as(hex.5km, "Spatial"), Hess = TRUE)
model3 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat), data = as(hex.5km, "Spatial"), Hess = TRUE)
model4 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem), as(hex.5km, "Spatial"), Hess = TRUE)
model5 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem) + scale(Rug), as(hex.5km, "Spatial"), Hess = TRUE)
model6 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rug), as(hex.5km, "Spatial"), Hess = TRUE)

@@

```{r}
##### 3.2. Include wildness components as additional predictors
model1 = polr(SQ ~ scale(Sce), data = as(hex.2km, "Spatial"), Hess = TRUE)
model2 = polr(SQ ~ scale(Sce) + scale(Abs), data = as(hex.2km, "Spatial"), Hess = TRUE)
model3 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat), data = as(hex.2km, "Spatial"), Hess = TRUE)
model4 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem), as(hex.2km, "Spatial"), Hess = TRUE)
model5 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem) + scale(Rug), as(hex.2km, "Spatial"), Hess = TRUE)
model6 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rug), as(hex.2km, "Spatial"), Hess = TRUE)

# GLM

```{r}
##### 3.1. Global Model: Generalised Linear Model
# hex.5km %>%
#   drop_na() %>%
#   as("Spatial") %>%
#   glm(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem) + scale(Rug),
#       data = .,
#       family = "poisson") %>%
#   summary()
# 
# R2 = 1 - glm$deviance/glm$null.deviance
```

# Machine learning - Gradient Boosted

```{r}
# GBM hyper-parameters tuning
# From kaggle: https://www.sciencedirect.com/science/article/pii/S0378778817320844
##### Split data into training and validation dataset
set.seed(2046)
TrainIndex <- hex.2km %>% drop_na() %>% .$SQ %>% createDataPartition(p = .8, list = FALSE, times = 1)
Train <- hex.2km[TrainIndex, ]
Test  <- hex.2km[-TrainIndex, ]

@@

# Reference
# https://github.com/topepo/caret/issues/389
```

```{r}
# Model evaluation
#predicted = predict(xgb_tune, x_test)
# Predict outcomes with the test data
Test %>% dplyr::select(Sce, Abs, Nat, Rem, Rug) %>%
  st_drop_geometry() %>% 
  as.matrix() %>% 
  xgb.DMatrix() %>%
  predict(xgb_tune, ., reshape = T) %>%
  plyr::mapvalues(from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")) %>%
  data.frame(predict = .) %>%
  data.frame(actual = plyr::mapvalues(Test$SQ, from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")),
             predict = .) ->
  result

@@

hex.2k$SQ.pred <-
  hex.2k %>%
  st_drop_geometry() %>%
  dplyr::select(Sce, Abs, Nat, Rem, Rug) %>%
  predict(xgb_tune, ., reshape = T) %>%
  plyr::mapvalues(from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")) %>%
  as.factor()
```

# Visualisation

```{r}
theme_map <- function(...) {
  theme_minimal() +
    theme(
      text = element_text(family = "Arial", color = "#22211d"), #"Ubuntu Regular"
      axis.line = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      # panel.grid.minor = element_line(color = "#ebebe5", size = 0.2),
      panel.grid.major = element_blank(), #element_line(color = "#ebebe5", size = 0.2),
      panel.grid.minor = element_blank(),
      plot.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA), 
      panel.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA), 
      legend.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA),
      panel.border = element_blank(),
      ...
    )
}
